### 2020-06-04 Outline.Rmd
Adding template for outline and list of packages

---
title: "Outline"
output: html_document
date: '`r format(Sys.time(), "%d %B, %Y")`'
---

<br/>

Journal: Current Landscape Ecology Reports
Type: Methodological Developments in Landscape Ecology

<br/>

< **TITLE** >

Maximillian H.K. Hesselbarth<sup>1</sup>, Jakub Nowosad<sup>2</sup>, ...

<br/>

<sup>1</sup>Department of Ecosystem Modelling, University of Goettingen, Buesgenweg 4, 37077 Goettingen, Germany

<sup>2</sup>Institute of Geoecology and Geoinformation, Adam Mickiewicz University, Krygowskiego 10, 61-680 Poznan, Poland

<br/>

*Corresponding author: E-Mail address: maximilian.hesselbarth@uni-goettingen.de

<br/>

Keywords: keyword 1, keyword 2, keyword 3, keyword 4, keyword 5, …

### 1. Introduction
- First topic
    - First paragraph headline
    - Second paragraph headline
    - ...
- Second topic
    - First paragraph headline
    - Second paragraph headline
    - ...
- ...
- Last topic (often objectives and hypotheses)
    - ...

### 2. Methods
- First topic
    - First paragraph headline
    - Second paragraph headline
    - ...
- Second topic
    - First paragraph headline
    - Second paragraph headline
    - ...
- ...
- Last topic
    - ...

### 3. Results (same structure and order as Methods)
- First topic
    - First paragraph headline
    - Second paragraph headline
    - ...
- Second topic
    - First paragraph headline
    - Second paragraph headline
    - ...
- ...
- Last topic (often objectives and hypotheses)
    - ...

### 4. Discussion
-	Possible points to discuss
    - First point
    - Second point
- ...
- Conclusions
    - Conclusion 1 
    - Conclusion 2
    - ...

### 5. Acknowledgments
- ...

### Open questions:
- ...
- ...