### 2021-05-03 Comments_Reviewer_R2.Rmd
Making changes to manuscript addressing R2

---
title: "Comments for the author R2"
output: pdf_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Comments for the Author

Thank you for your rapid revision and for the thoughtful way in which you addressed the comments. 
I am happy with what you have done to respond to the reviewers and I think it will be a valuable contribution to the journal. 
I have only a few minor edits - these are just minor wording/syntax issues and should not take long to fix.
Please see the attached marked-up PDF. 
It might also be nice to acknowledge the two anonymous reviewers in your acknowledgments section.

**Reply**: